### 2015-08-24 paper.tex
Add bare start to paper

\documentclass{article}

\title{SatNOGS: Satellite Networked Open Ground Station}
\author{Dan White}

\begin{document}

\maketitle

\begin{abstract}
asdfasdfasdf
asd
f
asdf
asdf
\end{abstract}

\section{Introduction}

\section{}

\end{document}