### 2015-08-24 paper.tex
use IEEE 2-column format

\documentclass[conference]{IEEEtran}

\usepackage[]{hyperref}

\author{
    \IEEEauthorblockN{Daniel J. White\IEEEauthorrefmark{1},
        SatNOGS Developers\IEEEauthorrefmark{2}}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}Valparaiso University, \href{mailto:dan.white@valpo.edu}{dan.white@valpo.edu}}
    \IEEEauthorblockA{\IEEEauthorrefmark{2}Libre Space Foundation,
    \href{mailto:info@satnogs.org}{info@satnogs.org}}

}

\title{SatNOGS: Satellite~Networked~Open~Ground~Station}

%TODO: add other authors

\begin{document}

\maketitle

\begin{abstract}
asdfasdfasdf
asd
f
asdf
asdf
\end{abstract}

\section{Introduction}
FOo

\section{Network}

\section{Database}

\section{Client}

\section{Ground Station}

\section{Conclusion}

\end{document}